Manuscript 2: Child Health Outcomes

### 2021-12-07 papertwo_child_health.Rmd
Update to switch to GEE for child health analysis

library(tidyverse)
library(gridExtra)
library(RColorBrewer)
library(tableone)
library(latex2exp)
library(labelled)
library(kableExtra)
library(logistf)
library(ARTool)
library(lme4)

@@

## Multivariable modeling results

@@

##watersource_model <- glmer(bottledwater ~ gender.x + age.x + racecollapse + pubben + (1 | unique_id), data = child_adult, family = "binomial") ##issues with convergence; did not include

##summary(watersource_model)

#For getting tested:
tested_model_noadult <- glmer(gettested ~ gender.x + age.x + racecollapse + pubben + (1 | unique_id), data = child_adult, family = "binomial")
tested_model_adult <- glmer(gettested ~ gender.x + age.x + racecollapse + pubben + water_exaggeratedbinary + (1 | unique_id), data = child_adult, family = "binomial")

summary(tested_model_noadult)
summary(tested_model_adult)

effects_mat <- data.frame("Outcome" = rep("Get tested", 6), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Water crisis was exaggerated (ref: Crisis not exaggerated)"), "Estimate" = c(summary(tested_model_noadult)$coefficients[-1,1], summary(tested_model_adult)$coefficients[7,1]), "StErr" = c(summary(tested_model_noadult)$coefficients[-1,2], summary(tested_model_adult)$coefficients[7,2]), "Pvalue" = c(summary(tested_model_noadult)$coefficients[-1,4], summary(tested_model_adult)$coefficients[7,4]))

#For having elevated BLLs:
#bll_model_noadult <- glmer(elevbll ~ gender.x + age.x + racecollapse + pubben + (1 | unique_id), data = child_adult, family = "binomial")
#bll_model_adult <- glmer(elevbll ~ gender.x + scale(age.x) + racecollapse + pubben + as.numeric(blood_test_after_april_2014=="Yes") + (1 | unique_id), data = child_adult, family = "binomial") # Not included due to convergence issues

#summary(bll_model_noadult)
#summary(bll_model_adult)

#subdat <- data.frame("Outcome" = rep("Elevated BLLs", 5), "Variable" = c("Gender", "Age", "Race-Other", "Race-White", "Public Benefits"), "Estimate" = c(summary(bll_model_noadult)$coefficients[-1,1]), "StErr" = c(summary(bll_model_noadult)$coefficients[-1,2]), "Pvalue" = c(summary(bll_model_noadult)$coefficients[-1,4]))

#effects_mat <- rbind(effects_mat, subdat)

#For having rashes:
rash_model_noadult <- glmer(rash ~ gender.x + age.x + racecollapse + pubben + (1 | unique_id), data = child_adult, family = "binomial")
summary(rash_model_noadult)
rash_model_adult <- glmer(rash ~ gender.x + age.x + racecollapse + pubben + as.numeric(blood_test_after_april_2014=="Yes") + (1 | unique_id), data = child_adult, family = "binomial")
summary(rash_model_adult)

subdat <- data.frame("Outcome" = rep("Rashes", 6), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Child screened for BLLs (ref: Child not screened for BLLs"), "Estimate" = c(summary(rash_model_noadult)$coefficients[-1,1], summary(rash_model_adult)$coefficients[7,1]), "StErr" = c(summary(rash_model_noadult)$coefficients[-1,2], summary(rash_model_adult)$coefficients[7,2]), "Pvalue" = c(summary(rash_model_noadult)$coefficients[-1,4], summary(rash_model_adult)$coefficients[7,4]))

@@

#For having hair loss:
#hair_model_noadult <- glmer(hair ~ gender.x + scale(age.x) + racecollapse + pubben + (1 | unique_id), data = child_adult, family = "binomial") ## convergence issues; did not include
#summary(hair_model_noadult)
#hair_model_adult <- glmer(hair ~ gender.x + scale(age.x) + racecollapse + pubben + as.numeric(blood_test_after_april_2014=="Yes") + (1 | unique_id), data = child_adult, family = "binomial") ## convergence issues; did not include
#summary(hair_model_adult)

#subdat <- data.frame("Outcome" = rep("Hairloss", 6), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Child screened for BLLs (ref: Child not screened for BLLs"), "Estimate" = c(summary(hair_model_noadult)$coefficients[-1,1], summary(hair_model_adult)$coefficients[7,1]), "StErr" = c(summary(hair_model_noadult)$coefficients[-1,2], summary(hair_model_adult)$coefficients[7,2]), "Pvalue" = c(summary(hair_model_noadult)$coefficients[-1,4], summary(hair_model_adult)$coefficients[7,4]))

#effects_mat <- rbind(effects_mat, subdat)

#For having learning delays:
learn_model_noadult <- glmer(learn ~ gender.x + age.x + racecollapse + pubben + (1 | unique_id), data = child_adult, family = "binomial")
summary(learn_model_noadult)

learn_model_layer2 <- glmer(learn ~ gender.x + age.x + racecollapse + pubben + as.numeric(blood_test_after_april_2014=="Yes") + (1 | unique_id), data = child_adult, family = "binomial")
summary(learn_model_layer2)

subdat <- data.frame("Outcome" = rep("Learning Delays", 6), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Child screened for BLLs (ref: Child not screened for BLLs"), "Estimate" = c(summary(learn_model_noadult)$coefficients[-1,1], summary(learn_model_layer2)$coefficients[7,1]), "StErr" = c(summary(learn_model_noadult)$coefficients[-1,2], summary(learn_model_layer2)$coefficients[7,2]), "Pvalue" = c(summary(learn_model_noadult)$coefficients[-1,4], summary(learn_model_layer2)$coefficients[7,4]))

@@

#For having hyperactivity:
hyper_model_noadult <- glmer(hyper ~ gender.x + scale(age.x) + racecollapse + pubben + (1 | unique_id), data = child_adult, family = "binomial")
summary(hyper_model_noadult)

hyper_model_layer2 <- glmer(hyper ~ gender.x + age.x + racecollapse + pubben + as.numeric(blood_test_after_april_2014=="Yes") + (1 | unique_id), data = child_adult, family = "binomial")
summary(hyper_model_layer2)

subdat <- data.frame("Outcome" = rep("Hyperactivity", 6), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Child screened for BLLs (ref: Child not screened for BLLs"), "Estimate" = c(summary(hyper_model_noadult)$coefficients[-1,1], summary(hyper_model_layer2)$coefficients[7,1]), "StErr" = c(summary(hyper_model_noadult)$coefficients[-1,2], summary(hyper_model_layer2)$coefficients[7,2]), "Pvalue" = c(summary(hyper_model_noadult)$coefficients[-1,4], summary(hyper_model_layer2)$coefficients[7,4]))

@@

#For having irritability:
irr_model_noadult <- glmer(irritable ~ gender.x + age.x + racecollapse + pubben + (1 | unique_id), data = child_adult, family = "binomial")
summary(irr_model_noadult)
irr_model_adult <- glmer(irritable ~ gender.x + age.x + racecollapse + pubben + as.numeric(blood_test_after_april_2014=="Yes") + (1 | unique_id), data = child_adult, family = "binomial")
summary(irr_model_adult)

subdat <- data.frame("Outcome" = rep("Irritability", 6), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Child screened for BLLs (ref: Child not screened for BLLs"), "Estimate" = c(summary(irr_model_noadult)$coefficients[-1,1], summary(irr_model_adult)$coefficients[7,1]), "StErr" = c(summary(irr_model_noadult)$coefficients[-1,2], summary(irr_model_adult)$coefficients[7,2]), "Pvalue" = c(summary(irr_model_noadult)$coefficients[-1,4], summary(irr_model_adult)$coefficients[7,4]))

@@

#For overall symptoms:
overall_model_noadult <- lmer(overall_symp ~ gender.x + age.x + racecollapse + pubben + (1 | unique_id), data = child_adult)
summary(overall_model_noadult)

overall_model_adult <- lmer(overall_symp ~ gender.x + age.x + racecollapse + pubben + as.numeric(blood_test_after_april_2014=="Yes") + water_exaggeratedbinary + (1 | unique_id), data = child_adult)
summary(overall_model_adult)

subdat <- data.frame("Outcome" = rep("Overall Symptoms", 7), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Child screened for BLLs (ref: Child not screened for BLLs)", "Water crisis was exaggerated (ref: Crisis not exaggerated)"), "Estimate" = c(summary(overall_model_noadult)$coefficients[-1,1], summary(overall_model_adult)$coefficients[7:8,1]), "StErr" = c(summary(overall_model_noadult)$coefficients[-1,2], summary(overall_model_adult)$coefficients[7:8,2]), "Pvalue" = pnorm(c(summary(overall_model_noadult)$coefficients[-1,3], summary(overall_model_adult)$coefficients[7:8,3]), lower.tail = FALSE))

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
effects_mat$BHpvalue <- p.adjust(effects_mat$Pvalue, method="BH")

effects_mat$lower <- case_when(
  effects_mat$Outcome=="Overall Symptoms" ~ effects_mat$Estimate - 1.96*effects_mat$StErr,
  effects_mat$Outcome!="Overall Symptoms" ~ exp(effects_mat$Estimate - 1.96*effects_mat$StErr)
)

effects_mat$upper <- case_when(
  effects_mat$Outcome=="Overall Symptoms" ~ effects_mat$Estimate + 1.96*effects_mat$StErr,
  effects_mat$Outcome!="Overall Symptoms" ~ exp(effects_mat$Estimate + 1.96*effects_mat$StErr)
)

effects_mat$effect <- case_when(
  effects_mat$Outcome=="Overall Symptoms" ~ effects_mat$Estimate,
  effects_mat$Outcome!="Overall Symptoms" ~ exp(effects_mat$Estimate)
)

effects_mat$label <- paste0(round(effects_mat$effect, digits=1), " (", round(effects_mat$lower, digits = 1), ", ", round(effects_mat$upper, digits = 1), ")")

pretty_effects <- select(effects_mat, Outcome, Variable, label, BHpvalue)
pretty_effects$BHpvalue <- round(pretty_effects$BHpvalue, 3)
rownames(pretty_effects) <- NULL

kable(pretty_effects, col.names = c("Outcome", "Predictor", "Odds Ratio (95% CI)", "P-value"))
```

Note that these modeling results doesn't include one important outcomes: child water source. I was unable to get this model to fit stably because of the small number of children not drinking bottled water. In addition, I couldn't consider the effect of the caregiver believing the water crisis was exaggerated on most of the outcomes because of convergence issues, and similarly with the relationship between the child being tested and the child having elevated BLLs.

>>>>

library(tidyverse)
library(gridExtra)
library(RColorBrewer)
library(tableone)
library(latex2exp)
library(labelled)
library(kableExtra)
library(logistf)
library(ARTool)
library(lme4)
library(rstanarm)
library(gee)

@@

## Multivariable modeling results (GEE)

@@

child_adult <- arrange(child_adult, unique_id)

#For getting tested:
tested_model <- gee((as.numeric(gettested)-1) ~ gender.x + age.x + racecollapse + pubben + water_exaggeratedbinary, id = unique_id, data = child_adult, family = "binomial", corstr = "exchangeable")

mysum <- summary(tested_model)$coefficients

mysum

effects_mat <- data.frame("Outcome" = rep("Get tested", 6), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Water crisis was exaggerated (ref: Crisis not exaggerated)"), "Estimate" = mysum[2:7, 1], "SE" = mysum[2:7, 4])

#For having elevated BLLs (restricting to children who were screened):
bll_model <- gee((as.numeric(elevbll)-1) ~ gender.x + age.x + racecollapse + pubben, id = unique_id, data = child_adult[child_adult$gettested=="Yes",], family = "binomial", corstr = "exchangeable")

mysum <- summary(bll_model)$coefficients

mysum

subdat <- data.frame("Outcome" = rep("Elevated BLLs", 5), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)"), "Estimate" = mysum[2:6, 1], "SE" = mysum[2:6, 4])

@@

#For having rashes:
rash_model <- gee((as.numeric(rash)-1) ~ gender.x + age.x + racecollapse + pubben + gettested  + elevbll, id = unique_id, data = child_adult, family = "binomial", corstr = "exchangeable")

mysum <- summary(rash_model)$coefficients

mysum

subdat <- data.frame("Outcome" = rep("Rashes", 7), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Child screened for BLLs (ref: Child not screened)", "Child diagnosed with elevated BLLs (ref: Child had normal BLLs)"), "Estimate" = mysum[2:8, 1], "SE" = mysum[2:8, 4])

@@

#For having learning delays:
learn_model <- gee((as.numeric(learn)-1) ~ gender.x + age.x + racecollapse + pubben + water_exaggeratedbinary + gettested + elevbll, id = unique_id, data = child_adult, family = "binomial", corstr = "exchangeable")

mysum <- summary(learn_model)$coefficients

mysum

subdat <- data.frame("Outcome" = rep("Learning delays", 8), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Water crisis was exaggerated (ref: Crisis not exaggerated)", "Child screened for BLLs (ref: Child not screened)", "Child diagnosed with elevated BLLs (ref: Child had normal BLLs)"), "Estimate" = mysum[2:9, 1], "SE" = mysum[2:9, 4])

@@

#For having hyperactivity:
hyper_model <- gee((as.numeric(hyper)-1) ~ gender.x + age.x + racecollapse + pubben + water_exaggeratedbinary + gettested + elevbll, id = unique_id, data = child_adult, family = "binomial", corstr = "exchangeable")

mysum <- summary(hyper_model)$coefficients

mysum

subdat <- data.frame("Outcome" = rep("Hyperactivity", 8), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Water crisis was exaggerated (ref: Crisis not exaggerated)", "Child screened for BLLs (ref: Child not screened)", "Child diagnosed with elevated BLLs (ref: Child had normal BLLs)"), "Estimate" = mysum[2:9, 1], "SE" = mysum[2:9, 4])

@@

#For having irritability:
irr_model <- gee((as.numeric(irritable)-1) ~ gender.x + age.x + racecollapse + pubben + water_exaggeratedbinary + gettested + elevbll, id = unique_id, data = child_adult, family = "binomial", corstr = "exchangeable")

mysum <- summary(irr_model)$coefficients

mysum

subdat <- data.frame("Outcome" = rep("Irritability", 8), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Water crisis was exaggerated (ref: Crisis not exaggerated)", "Child screened for BLLs (ref: Child not screened)", "Child diagnosed with elevated BLLs (ref: Child had normal BLLs)"), "Estimate" = mysum[2:9, 1], "SE" = mysum[2:9, 4])

#For overall symptoms:
overall_model <- gee(overall_symp ~ gender.x + age.x + racecollapse + pubben + water_exaggeratedbinary + gettested + elevbll, id = unique_id, data = child_adult, family = "poisson", corstr = "exchangeable") 
mysum <- summary(overall_model)$coefficients

mysum

subdat <- data.frame("Outcome" = rep("Total Symptoms", 8), "Variable" = c("Male gender (ref: Female gender)", "Age", "Other race (ref: Black race)", "White race (ref: Black race)", "Receives public benefits (ref: No benefits)", "Water crisis was exaggerated (ref: Crisis not exaggerated)", "Child screened for BLLs (ref: Child not screened)", "Child diagnosed with elevated BLLs (ref: Child had normal BLLs)"), "Estimate" = mysum[2:9, 1], "SE" = mysum[2:9, 4])

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
effects_mat$Lower <- exp(effects_mat$Estimate - 1.96*effects_mat$SE)
effects_mat$Upper <- exp(effects_mat$Estimate + 1.96*effects_mat$SE)
effects_mat$OR <- exp(effects_mat$Estimate)
effects_mat$pvalue <- pnorm(abs(effects_mat$Estimate/effects_mat$SE), lower.tail = FALSE)
effects_mat$BHPvalue <- p.adjust(effects_mat$pvalue, method="BH")

effects_mat$prettyp <- round(effects_mat$pvalue, digits = 2)
effects_mat$prettypBH <- round(effects_mat$BHPvalue, digits = 2)
effects_mat$sig <- effects_mat$pvalue < 0.05

effects_mat$label <- paste0(round(effects_mat$OR, digits=1), " (", round(effects_mat$Lower, digits = 1), ", ", round(effects_mat$Upper, digits = 1), ")")

pretty_effects <- dplyr::select(effects_mat, Outcome, Variable, label, prettypBH)
rownames(pretty_effects) <- NULL

kable(pretty_effects)
#kable(pretty_effects, col.names = c("Outcome", "Predictor", "Odds Ratio (95% CI)"))
```
